Solving the power flow problem on integrated transmission-distribution networks: a review and numerical assessment

### 2020-02-06 IEEEtran/paper.tex
results tables finished. only text

\newpage 
\section{Numerical assessment}
All previous mentioned methods are implemented into the Matpower\footnote{MATPOWER is a package of free, open-source Matlab-language M-files for solving steady-state power system simulation and optimization problems \cite{Zimmerman2011}} library. Matpower consists of several transmission network test cases. The resources page of IEEE Power \& Energy Society  contains several distribution network test cases which are all explained in \cite{Schneider2018}. These test cases contain the necessary input to solve power flow problems. We created integrated test test cases from the existing transmission and distribution test cases. We used the IEEE 9-bus, 118-bus, and 3120-bus data as transmission network test cases and the IEEE 13-bus and 37-bus data as distribution test cases. The distribution test cases represent the unbalanced characteristics of the network. We modified the 13-bus distribution network to a 10-bus network by deleting the buses connected to the regulator. We changed the loading of the 37-bus network by shifting 20\% of the loads of phase b equally to phase a and c, just like the original authors \cite{Taranto2008}.
We connected two networks to each other creating the following integrated test cases: 
\begin{itemize}
    \item Test case 1: T9-D13
    \item Test case 2: T9-D37
    \item Test case 3: T118-D37
    \item Test case 4: T3120-D37
    \item Test case 5: T9-2D13 (2 13-bus distribution networks) 
    \item Test case 6: T9-3D13 (3 13-bus distribution networks) 
\end{itemize}
\paragraph{Connection bus}
We selected a random load bus in the transmission network to become the connection bus in the integrated network. We choose bus $7$ in the 9-bus network, we choose bus $118$ in the 118-bus network, and we choose bus $3003$ in the 3120-bus network. This could have been any load bus of the original network. The original reference bus of the distribution network becomes the connection bus in the integrated network. \\
The unified methods puts a transformer between the connection buses. This transformer has the same line admittance as the first distribution cable. The lhs of the transformer is the transmission connection bus, now having zero load. The rhs is the distribution connection bus. This is the former reference bus and must be changed to a load bus having the load as the original transmission connection bus \cite{Taranto2008}. The splitting methods do not put a transformer between the networks. They only change the original transmission load bus to a bus having zero load.   \newline \newline
Test case 5 and 6 have multiple distribution networks connected. These networks are connected to the last 2 and 3 load buses of the transmission networks, where each of the original load bus is changed to a zero load bus. 
\subsection{Results}

\begin{table}[!ht]
\renewcommand{\arraystretch}{1.3}
\centering
\label{tab:resultsPIC3P}\hspace{2cm}
\caption{Comparison on number of iterations for homogeneous networks (for the MSS-method ($I_{MSS}$) and the necessary iterations per system ($I_T$ and $I_D$)), and CPU-time of the four different MSS-methods, applied on five test-cases. The numbers marked with an asterisk converged to wrong results. In this case, the MAI method required more than 2 subiterations to converge correctly. The slowest CPU times are printed in bold. The top one are methods applied on hybrid networks. The bottom one is applied on homogeneous networks. }
\subfloat[Hybrid networks]{
\begin{adjustbox}{width=1\textwidth} %, angle=90}
\small
\begin{tabular}{@{}l c cc c  cccc c cccc c  @{}}\toprule
                               && \multicolumn{2}{c}{\textit{IC}} &&     \multicolumn{4}{c}{\textit{MS-hybrid-CAI}} && \multicolumn{4}{c}{\textit{MS-hybrid-MAI}} \\ \toprule 
\multicolumn{1}{l}{}        && \textit{its}      & \textit{CPU} && $I_{MSS}$      & $I_T$   &  $I_D$      & \textit{CPU}     &&$I_{MSS}$      & $I_T$   &  $I_D$      & \textit{CPU}      \\
\cmidrule{3-4}  \cmidrule{6-9}  \cmidrule{11-14}   
\multicolumn{1}{l}{}      && \textit{\#}       & \textit{sec} && \textit{\#}      & \textit{\#}    & \textit{\#}       & \textit{sec}     && \textit{\#}        & \textit{\#}     &  \textit{\#}       & \textit{sec}  \\
\midrule
\multicolumn{1}{l}{\textit{T9-D13 (7)}}         && 4 & {0.455 } && 4       & 4  & 4 & 0.308 && 6      & 2  & 2 & 0.323  \\
\multicolumn{1}{l}{\textit{T9-D37 (7)}}         && 4 & 0.464    && 3      & 4  & 4 & 0.325 && 6      & 2  & 2 & 0.352  \\
\multicolumn{1}{l}{{\textit{T118-D13 (89)}}}    && 4 & 0.522    && 3      & 7 & 5 & 0.348 && 4 & 2  & 2 & {0.312*} \\
\multicolumn{1}{l}{{\textit{T118-D37 (89)}}}    && 4 & 0.518    && 3      & 7 & 5 & 0.356 && 4 & 2  & 2 & {0.358}\\
\multicolumn{1}{l}{{\textit{T3120-D37 (3003)}}} && 5 & 143      && 3      & 6  & 4  & 383 && 6*  & 2*  & 2* & {235*}  \\
% \multicolumn{1}{l}{\textit{T9-2D13}}       && - & - && - & -    && -       & - & -  & \textbf{-} && -   & - & - &  - && -       & -  & - & - && -       & -  & - & - \\
% \multicolumn{1}{l}{{\textit{T9-3D13}}}    && - & - && - & -    && -       & - & -  & \textbf{-} && -   & - & - &  - && -       & -  & - & - && -       & -  & - & - \\
\toprule 
\end{tabular}
\end{adjustbox}}

\hspace{-2.5cm}
\subfloat[Homogeneous networks]{
%\caption{.}
\label{tab:resultsPIC3P}\hspace{2cm}
\begin{adjustbox}{width=1\textwidth} %, angle=90}
\small
\begin{tabular}{@{}l c cc c  cccc c cccc c  @{}}\toprule
                               && \multicolumn{2}{c}{\textit{F3P}} &&     \multicolumn{4}{c}{\textit{MS-homo-CAI}} && \multicolumn{4}{c}{\textit{MS-homo-MAI}} \\ \toprule 
\multicolumn{1}{l}{}        && \textit{its}      & \textit{CPU} && $I_{MSS}$      & $I_T$   &  $I_D$      & \textit{CPU}     &&$I_{MSS}$      & $I_T$   &  $I_D$      & \textit{CPU}      \\
\cmidrule{3-4}  \cmidrule{6-9}  \cmidrule{11-14}   
\multicolumn{1}{l}{}      && \textit{\#}       & \textit{sec} && \textit{\#}      & \textit{\#}    & \textit{\#}       & \textit{sec}     && \textit{\#}        & \textit{\#}     &  \textit{\#}       & \textit{sec}  \\
\midrule
\multicolumn{1}{l}{\textit{T9-D13 (7)}}      && 4 &\textbf{ 0.610}  && 4 & 4 & 4 &  0.237 && 6       & 2 & 2  & {0.257}\\
\multicolumn{1}{l}{\textit{T9-D37 (7)}}      && 4 & \textbf{0.587}    && 3 & 4 & 4 &  0.262 && 6       & 2 & 2  & {0.299}\\
\multicolumn{1}{l}{{\textit{T118-D13 (89)}}}   && 5 & \textbf{0.763}     && 3       & 7 & 5  & 0.324 && 4  & 2 &2 &  0.297  \\
\multicolumn{1}{l}{{\textit{T118-D37 (89)}}}   && 5 & \textbf{0.920}    && 3       & 7 & 5  & 0.360 && 4 & 2 & 2 &  0.325 \\
\multicolumn{1}{l}{{\textit{T3120-D37 (3003)}}}   && 5 & \textbf{2438}   &&  3 & 6  & 4  & 1213  && 6 & 2 & 2 &  841    \\
% \multicolumn{1}{l}{\textit{T9-2D13}}       && - & - && - & -    && -       & - & -  & \textbf{-} && -   & - & - &  - && -       & -  & - & - && -       & -  & - & - \\
% \multicolumn{1}{l}{{\textit{T9-3D13}}}    && - & - && - & -    && -       & - & -  & \textbf{-} && -   & - & - &  - && -       & -  & - & - && -       & -  & - & - \\
\toprule 
\end{tabular}
\end{adjustbox}}

\begin{table}[!ht]
\renewcommand{\arraystretch}{1.3}
\centering
\label{tab:resultsPIC3P}\hspace{2cm}
\begin{adjustbox}{width=1\textwidth} %, angle=90}
\small
\begin{tabular}{@{}l c cc c  cccc c cccc c  @{}}\toprule
                               && \multicolumn{2}{c}{\textit{F3P}} &&     \multicolumn{4}{c}{\textit{MS-homo-CAI}} && \multicolumn{4}{c}{\textit{MS-homo-MAI}} \\ \toprule 
\multicolumn{1}{l}{}        && \textit{its}      & \textit{CPU} && $I_{MSS}$      & $I_T$   &  $I_D$      & \textit{CPU}     &&$I_{MSS}$      & $I_T$   &  $I_D$      & \textit{CPU}      \\
\cmidrule{3-4}  \cmidrule{6-9}  \cmidrule{11-14}   
\multicolumn{1}{l}{}      && \textit{\#}       & \textit{sec} && \textit{\#}      & \textit{\#}    & \textit{\#}       & \textit{sec}     && \textit{\#}        & \textit{\#}     &  \textit{\#}       & \textit{sec}  \\
\midrule
\multicolumn{1}{l}{\textit{T9-D13 (7)}}      && 4 &\textbf{ 0.610}  && 4 & 4 & 4 &  0.237 && 6       & 2 & 2  & \textbf{0.257}\\
\multicolumn{1}{l}{\textit{T9-D37 (7)}}      && 4 & \textbf{0.587}    && 3 & 4 & 4 &  0.262 && 6       & 2 & 2  & \textbf{0.299}\\
\multicolumn{1}{l}{{\textit{T118-D13 (89)}}}   && 5 & \textbf{0.763}     && 3       & 7 & 5  & 0.324 && 4  & 2 &2 &  0.297  \\
\multicolumn{1}{l}{{\textit{T118-D37 (89)}}}   && 5 & \textbf{0.920}    && 3       & 7 & 5  & 0.360 && 4 & 2 & 2 &  0.325 \\
\multicolumn{1}{l}{{\textit{T3120-D37 (3003)}}}   && 5 & \textbf{2438}   &&  3 & 6  & 4  & 1213  && 6 & 2 & 2 &  841    \\
% \multicolumn{1}{l}{\textit{T9-2D13}}       && - & - && - & -    && -       & - & -  & \textbf{-} && -   & - & - &  - && -       & -  & - & - && -       & -  & - & - \\
% \multicolumn{1}{l}{{\textit{T9-3D13}}}    && - & - && - & -    && -       & - & -  & \textbf{-} && -   & - & - &  - && -       & -  & - & - && -       & -  & - & - \\
\toprule 
\end{tabular}
\end{adjustbox}
\caption{Comparison on number of iterations for homogeneous networks (for the MSS-method ($I_{MSS}$) and the necessary iterations per system ($I_T$ and $I_D$)), and CPU-time of the four different MSS-methods, applied on five test-cases. The numbers marked with an asterisk converged to wrong results. In this case, the MAI method required more than 2 subiterations to converge correctly. The slowest CPU times are printed in bold.}
\end{table}
\newpage
\input{L-MAIN2}

\input{L-MAIN1}
\input{F3Pnorms}
\input{ICnorms}

\input{L-N2I1}
\input{L-N1I1}
\input{L-N2I2}
\input{L-N1I2}
\newpage 
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Images/MAI_N1.eps}
\caption{Hybrid networks}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Images/Z_MAI_N1.eps}
\caption{Hybrid networks}
\end{figure}
\newpage
The results show that most of the methods are good options to solve integrated networks. They converge in a reasonable amount of time and number of iterations and this number is not increasing when the size of the network increases. Only the MS-homogeneous-MAI method is not working properly. The method does not reach convergence. If we take a closer look at the voltage profiles, Figure ... , we see a difference in how this is distributed according to the separated networks. We start with the unified methods. We see a drop at bus $7$, this is the bus connected to the distribution network via a transformer. In the Master-Slave methods we see a different profile at bus $7$: the voltage here is equal to the voltage at the first bus of the distribution network (bus 10 in the plot). But bus $2$ is significantly lower. The results differ, because the MSS-algorithm does not put a transformer in between the two connecting nodes: it uses the input and output directly. So bus $7$ IS bus $1$ of the distribution network. We then must see the voltage drop somewhere else in the transmission network, which we do: mainly at bus 2. This bus is directly connected to bus 7, so now gets less power. \\
Distribution networks are radial, so power flows into the first node and is then distributed to the rest of the network. So all these buses are only influenced by the previous bus. That is why the voltage profiles of distribution networks remains the same, only the voltage magnitude of the entire network is higher, because they receive more power from the connected transmission network, then they would get from the generator when the network is modeled separately. 
\begin{figure*}[h!]
    \centering
    \includegraphics[width=\textwidth]{Images/Enumathpaper.eps}
    \caption{Voltage magnitudes of the different integration methods and separated method. Up: The MSS-methods. Down: The unified methods. Bus 7 and 11 are the buses that are connected.}
    \label{fig:voltmag}
\end{figure*}
\subsection{Speeding-up the computations}
Real transmission and distribution networks, in the Netherlands, have millions of buses. The time to run simulations then becomes really important. In the test-simulations, it is difficult to draw any conclusions from the CPU-time. Only the T2383-D13 network gives a visible difference in CPU-time. As expected, the F3P-method is somewhat slower than the IC-method, and the MSS-methods are even slower. This is logical, because in the F3P-method, the size of the Jacobian increases of being 2n by 2n, to 6n by 6n, n the number of buses in the network. The MSS-method requires 4 iterations, thus solving each the transmission network and distribution network 4 times, instead of one integrated network 4 times. \\
But real networks have another difference, compared to the T2383 network: real distribution networks are much larger than real transmission networks. So the size of transmission networks is not really influencing the speed of computations. Furthermore, very large networks, must be solved in parallel and the MSS-method is some kind of domain-decomposition approach: multiple distribution networks can run on parallel cores and then inject the result to the main core, where the transmission network is solved. Doing new simulations using these speed-up techniques on realistically sized networks, should give a better idea which method is most favourable to solve the integrated power flow problem. 
\section{Discussion}

>>>>

\newpage 
\section{Numerical assessment}
All previous mentioned methods are implemented into the Matpower\footnote{MATPOWER is a package of free, open-source Matlab-language M-files for solving steady-state power system simulation and optimization problems \cite{Zimmerman2011}} library. Matpower consists of several transmission network test cases. The resources page of IEEE Power \& Energy Society  contains several distribution network test cases which are all explained in \cite{Schneider2018}. These test cases contain the necessary input to solve power flow problems. We created integrated test test cases from the existing transmission and distribution test cases. We used the IEEE 9-bus, 118-bus, and 3120-bus data as transmission network test cases and the IEEE 13-bus and 37-bus data as distribution test cases. The distribution test cases represent the unbalanced characteristics of the network. We modified the 13-bus distribution network to a 10-bus network by deleting the buses connected to the regulator. We changed the loading of the 37-bus network by shifting 20\% of the loads of phase b equally to phase a and c, just like the original authors \cite{Taranto2008}.
We connected two networks to each other creating the following integrated test cases: 
\begin{itemize}
    \item Test case 1: T9-D13
    \item Test case 2: T9-D37
    \item Test case 3: T118-D37
    \item Test case 4: T3120-D37
    \item Test case 5: T9-2D13 (2 13-bus distribution networks) 
    \item Test case 6: T9-3D13 (3 13-bus distribution networks) 
\end{itemize}
\paragraph{Connection bus}
We selected a random load bus in the transmission network to become the connection bus in the integrated network. We choose bus $7$ in the 9-bus network, we choose bus $118$ in the 118-bus network, and we choose bus $3003$ in the 3120-bus network. This could have been any load bus of the original network. The original reference bus of the distribution network becomes the connection bus in the integrated network. \\
The unified methods puts a transformer between the connection buses. This transformer has the same line admittance as the first distribution cable. The left hand side of the transformer is the transmission connection bus, now having zero load. The right hand side is the distribution connection bus. This is the former reference bus and must be changed to a load bus having the load as the original transmission connection bus \cite{Taranto2008}. The splitting methods do not put a transformer between the networks. They only change the original transmission load bus to a bus having zero load.   \newline \newline
Test case 5 and 6 have multiple distribution networks connected. These networks are connected to the last 2 and 3 load buses of the transmission networks, where each of the original load bus is changed to a zero load bus. 
\subsection{Results}
We are going to compare the above-mentioned methods on accuracy, speed, and convergence to give an insight into which method is favorable to solve integrated networks. 
\subsubsection{Accuracy} 
To check the accuracy of the methods, we compare the voltage output of the integrated methods with the output of the separated methods. We solve the transmission network separately using NR-power and we solve the distribution network separately using NR-TCIM. The distribution domain in the integrated network has no reference bus anymore. In order to compare the voltage profile of this part, we scale the values accordingly. \\
We know that the voltage is given by: \begin{equation}
V_p=|V|_p\exp{(\iota\delta_{V_p}-\phi)},\quad p\in\{a,b,c\},
\end{equation}
where $|V|$ and $\delta$ are set to $|V|=1.0$ pu and $\delta=0$ at the reference bus. We can compare the voltage of the distribution domain to the voltage of the separated domain after scaling. We divide the voltage magnitude of the distribution domain, buses $i=1,...,D_N$ by the voltage magnitude of the distribution connection bus: 
\begin{equation}
|V|_{i,{D_{new}}} = \frac{|V|_{i,{D_{uns}}}}{|V|_{ref,{D_{uns}}}},\quad i=1,..,D_N.
\label{eq:ViDnew}
\end{equation}
The voltage angle represents the phase-shift in relation to the reference bus. In order to compare the integrated output of the voltage angle with the separated output, we subtract the phase-angle of the former reference bus of all the buses of the distribution domain: 
\begin{equation}
\delta_{i,{D_{new}}} ={\delta_{i,{D_{uns}}}}-{\delta_{ref,{D_{uns}}}},\quad i=1,..,D_N.
\label{eq:dViDnew}
\end{equation}
Note that we need to apply \eqref{eq:ViDnew} and \eqref{eq:dViDnew} to all the three phases separately. \newline\newline
We compare the voltage magnitudes and angles of the integrated network, including the scaled distribution domain, with the magnitudes angles of the separate domains. We take the infinity norm of the relative difference between the voltages: 
\begin{equation}
\mbox{rel. error of } |V|_p = \left|\frac{|V|^*_p - |V|^o_p}{|V|^o_p}\right|_\infty,\quad p\in\{a,b,c\},
\label{eq:norm|V|}
\end{equation}
where the asterisk $*$ marks the outcome of separated networks, and the superscript $o$ the output of separated networks (original). We do this in a similar manner for the voltage angle: 
\begin{equation}
\mbox{rel. error of } \delta_p = \left|\frac{\delta^*_p - \delta^o_p}{\delta^o_p}\right|_\infty,\quad p\in\{a,b,c\},
\label{eq:normdelta}
\end{equation}
We have given the relative errors of the voltage magnitude and angle of the three phases for all the test cases in table \ref{tab:relerror}.

\input{allnorms}
\paragraph{Relative error}
Table \ref{tab:relerror} shows that all the integration methods produce accurate results, except for the MSS-homo-MAI and MSS-hybrid-MAI methods. This is caused by the number of iterations per subdomain that has been set too strict. We increased the number of iterations per subdomain to see whether the output gets more accurate. We show these results in table \ref{tab:MAIschems}. Next to the relative error, this table shows the required MSS iterations and how long it takes to reach convergence. We only applied these tests on test case 4, as these results were least accurate.

\input{L-MAIN2}

%\input{L-MAIN1}

\subsubsection{Convergence and speed} 
The previous subsection shows that all the methods produce accurate results. In order to solve realistic power flow problems, which are very large electricity networks, we need insight into the speed of the problems. In this section we compare the integration methods on CPU-time and number of iterations. We share the results in table \ref{tab:speed}. 
\include{tablespeed}

% Nu de figuren

\newpage 
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Images/MAI_N1.eps}
\caption{Hybrid networks}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Images/Z_MAI_N1.eps}
\caption{Hybrid networks}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Images/MAI_N2.eps}
\caption{Homogeneous networks}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Images/Z_MAI_N2.eps}
\caption{Homogeneous networks}
\end{figure}

\newpage
The results show that most of the methods are good options to solve integrated networks. They converge in a reasonable amount of time and number of iterations and this number is not increasing when the size of the network increases. Only the MS-homogeneous-MAI method is not working properly. The method does not reach convergence. If we take a closer look at the voltage profiles, Figure ... , we see a difference in how this is distributed according to the separated networks. We start with the unified methods. We see a drop at bus $7$, this is the bus connected to the distribution network via a transformer. In the Master-Slave methods we see a different profile at bus $7$: the voltage here is equal to the voltage at the first bus of the distribution network (bus 10 in the plot). But bus $2$ is significantly lower. The results differ, because the MSS-algorithm does not put a transformer in between the two connecting nodes: it uses the input and output directly. So bus $7$ IS bus $1$ of the distribution network. We then must see the voltage drop somewhere else in the transmission network, which we do: mainly at bus 2. This bus is directly connected to bus 7, so now gets less power. \\
Distribution networks are radial, so power flows into the first node and is then distributed to the rest of the network. So all these buses are only influenced by the previous bus. That is why the voltage profiles of distribution networks remains the same, only the voltage magnitude of the entire network is higher, because they receive more power from the connected transmission network, then they would get from the generator when the network is modeled separately. 
\begin{figure*}[h!]
    \centering
    \includegraphics[width=\textwidth]{Images/Enumathpaper.eps}
    \caption{Voltage magnitudes of the different integration methods and separated method. Up: The MSS-methods. Down: The unified methods. Bus 7 and 11 are the buses that are connected.}
    \label{fig:voltmag}
\end{figure*}
\subsection{Speeding-up the computations}
Real transmission and distribution networks, in the Netherlands, have millions of buses. The time to run simulations then becomes really important. In the test-simulations, it is difficult to draw any conclusions from the CPU-time. Only the T2383-D13 network gives a visible difference in CPU-time. As expected, the F3P-method is somewhat slower than the IC-method, and the MSS-methods are even slower. This is logical, because in the F3P-method, the size of the Jacobian increases of being 2n by 2n, to 6n by 6n, n the number of buses in the network. The MSS-method requires 4 iterations, thus solving each the transmission network and distribution network 4 times, instead of one integrated network 4 times. \\
But real networks have another difference, compared to the T2383 network: real distribution networks are much larger than real transmission networks. So the size of transmission networks is not really influencing the speed of computations. Furthermore, very large networks, must be solved in parallel and the MSS-method is some kind of domain-decomposition approach: multiple distribution networks can run on parallel cores and then inject the result to the main core, where the transmission network is solved. Doing new simulations using these speed-up techniques on realistically sized networks, should give a better idea which method is most favourable to solve the integrated power flow problem. 
\section{Discussion}
